---
title: "Homework1"
author: "Peter Kamala - 20210818"
format: 
  revealjs:
    theme: solarized
    slide-number: true
    transition: zoom
execute: 
  eval: true
  echo: false
  message: false
  warning: false
fontsize: 14pt
---

## [Introduction]{style="color: blue;"}

In this presentation I'll examine the <strong>UK Quarterly Gas Consumption</strong> dataset, containing UK gas consumption from the first quarter of 1960 to the fourth quarter of 1986. <br>

The dataset has four quarterly record of gas consumption in millions of therms for each year from 1960 to 1986 as seen in the table below <br><br>

```{r}
library(knitr)
library(datasets)
library(tidyr)
library(DT)
library(ggplot2)

# Create dataframe
df_ukg <- data.frame(
  Year = floor(time(UKgas)),
  Quarter = cycle(UKgas),
  GasConsumption = as.numeric(UKgas)
)

# Pivot the quarters
ukgas_wider <- pivot_wider(df_ukg, names_from = Quarter, values_from = GasConsumption)

# Display the pivoted table
colnames(ukgas_wider) <- c("Year", "Q1", "Q2", "Q3", "Q4")
datatable(ukgas_wider, caption = 'UK Quarterly Gas Consumption by Year (1960-1986)', options = list(pageLength = 9))

```

## [Descriptive Statistics for the quartery gas consumption variable]{style="color: blue;"}

```{r}
summary(ukgas_wider[, -1])
```

As the statistics indicate, quarter 1 has the highest recorded gas consumption, while quarter 3 has the lowest. Could be due the cold months of quarter 1 and the opposite on quarter 3.

## [Plotting the UK Quarterly Gas consumption dataset]{style="color: blue;"}

```{r}
# Melting the ukgas_wider dataset for plotting
ukgas_long <- pivot_longer(ukgas_wider, cols = -Year, names_to = "Quarter", values_to = "GasConsumption")

# Plotting the dataset to show the seasonal(quarterly) patterns more clearly.
ggplot(ukgas_long, aes(x = Year, y = GasConsumption, color = Quarter)) +
  geom_line() +
  labs(x = "Year", y = "Gas Consumption (millions of therms)",
       title = "UK Quarterly Gas Consumption (1960-1986)")
```

The above plot visualizes UK quarterly gas consumption from <b>1960</b> to <b>1986</b>, relatively higher consumption is observed during quarter 1 across all the years, could be due to being a cold season, also relatively low consumption is observed on quarter 3, could be due to being a warm season.

```{r}
usethis::git_sitrep()
```

```{r}
 usethis::create_github_token()
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
```

```{r}
 gitcreds::gitcreds_set()
```